---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# eavtools

The goal of eavtools is to provide utility functions for working with data in Entity-Attribute-Value format *as if* it were tidy.

## Installation

You can install the development version of this package from github with:

``` r
# install.packages("devtools")
devtools::install_github("jameelalsalam/eavtools")
```

## Example

Entity-attribute value can be a convenient format for storing many different variables about model aspect. Using the key-value pair of columns `variable` and `value` is the marker for entity-attribute-value formatted data:

```{r example}
## basic example code
library(tidyverse)
library(eavtools)

eav_emissions
```


One common need for data in this format is to perform calculations across a group of rows. For example, what does each model project for the change in emissions from 2015 to 2030?


```{r}
eav_emissions %>%
	filter(model!= "m3") %>%
	group_by(model) %>%
	summarize(emissions_change = value[variable == "emissions" & year == 2030] / value[variable == "emissions" & year == 2015])
```

There are a couple of problems here:

- I don't find it easy to read.
- This type of code doesn't handle missing or duplicate values well. If a value is _implicitly_ missing (e.g., there is no row corresponding to the selected years/variables) then the code will throw an error. Summarize needs something length 1, not length 0.

```{r, error=TRUE}
eav_emissions %>%
	filter(model== "m3") %>%
	group_by(model) %>%
	summarize(emissions_change = value[variable == "emissions" & year == 2030] / value[variable == "emissions" & year == 2015])
```

```{r}
eav_emissions %>%
	group_by(model) %>%
	summarize(emissions_change = 
							value %forwhich% (variable == "emissions" & year == 2030) /
							value %forwhich% (variable == "emissions" & year == 2015))
```

Using the `%forwhich%` function in this package solves the error behavior, but it does not solve the readability issue. This type of code scales up:

```{r}
eav_emissions %>%
	group_by(model, variable) %>%
	summarize(change_2015_to_2030 = 
							value %forwhich% (year == 2030) /
							value %forwhich% (year == 2015))
```

